\tool: Revisiting BadUSB with Type-C

### 2021-03-22 WOOT/Files/background.tex

We first introduce the development of \ac{USB} specification and emphasize the key
points adopted in this work. We also organized a brief timeline for introducing
key points of each protocol in Table~\ref{table:usb_timeline}.

@@

%\noindent\outline{HID Protocol}\\
Right after one year of the appearance of \ac{USB} 1.0, a standard named \acf{HID}~\cite{hid} was designed based on \ac{USB}. \ac{HID} is
designed to unify the implementation for devices like mice,
keyboards, etc. Before its appearance, the standard is divided among
manufacturers, for example, the mouse of Company A may use X-Y coordinates to
represent its location while the mouse of Company B uses relative displacement.
This means every device needs its own driver to work. After \ac{HID}, users only need to
write one driver for an entire class of \acp{HID}. Furthermore, the \ac{HID} standard also
requires all devices to be \ac{PnP}, which is indeed
convenient but insecure too.

In 1998, the first widely supported \ac{USB} protocol was designed. \ac{USB} 1.1~\cite{usb11}
provided two data transfer rates which are low speed (1.5 Mbit/s) and full
speed (12 MBit/s). At this point, due to the transfer limitation, it only supports
limited types of devices like keyboards, mice, etc.

%\noindent\outline{USB2.0}\\
In 2000, the \ac{USB} 2.0~\cite{usb20} specification was released. With high speed \mbox{(480
Mbit/s)} mode introduced, printers, cameras, CD-ROM drives, and network cards were
supported in this revision. Such a high data transfer rate also gave rise to the
popularity of ``flash drive'', a portable device that allows physically
transferring data around~\cite{sok}. Although various peripherals were supported
in \ac{USB} 2.0, there was no reliable way to identify the type of device. This
security flaw allowed attacks like BadUSB~\cite{badusb,rubber}.

%\noindent\outline{USB3.x}\\
\ac{USB} 3.0~\cite{usb30} was introduced in 2008, with a super speed \mbox{(5 Gbit/s)} data
transfer rate. Like its predecessor, more classes of peripherals were supported
in this revision. In 2013, the \ac{USB} Type-C connector standard was introduced as a
part of \ac{USB} 3.1~\cite{usb31}, providing a unified connector type for
PowerDelivery, Thunderbolt, DisplayPort, and HDMI.  Yet no improvement of
security was introduced in 3.x revisions, meaning any device claiming
itself as a monitor can capture the video stream from the host. Exposing
such a multi-purpose connector unprotected is insecure and allows attacks similar to \tool. In 2017, \ac{USB} 3.2~\cite{usb32} was released, doubling the data
transfer rate (20 Gbit/s).

@@

Thus, to provide support towards a wider range of peripherals, a 24-pins
standard called \ac{USB} Type-C~\cite{typec} was introduced in 2013 by \ac{USB}-IF~\cite{usbif}. As it is designed to be double-sided, the number of actually usable
pins is halved. Nevertheless, this standard has largely enhanced the capability
of the \ac{USB} 3.x protocol. As presented in Figure~\ref{fig:usb_c_conn}, Type-C added
two high-speed data lanes \mbox{(TRX1 \& TRX2)} and kept the original data lane \mbox{(D+ \&
D-)}. The added lanes are used exclusively to support peripherals like
DisplayPort while the kept data lane transfers \ac{USB} packets.

>>>>

We first introduce the development of \ac{USB} specification and emphasize the key
points adopted in this work. We also organize a brief timeline for introducing
key points of each protocol in Table~\ref{table:usb_timeline}.

@@

%\noindent\outline{HID Protocol}\\
Right after one year of the appearance of \ac{USB} 1.0, a standard named \acf{HID}~\cite{hid} was designed based on \ac{USB}. \ac{HID} is
designed to unify the implementation for devices like keyboards,
mice, etc. Before its appearance, the standard is divided among
manufacturers, for example, the mouse of Company A may use X-Y coordinates to
represent its location while the mouse of Company B uses relative displacement.
This means every device needs its own driver to work. After \ac{HID}, users only need to
write one driver for an entire class of \acp{HID}. Furthermore, the \ac{HID} standard also
requires all devices to be \ac{PnP}, which is indeed
convenient but insecure too.

In 1998, the first widely supported \ac{USB} protocol was designed. \ac{USB} 1.1~\cite{usb11}
provides two data transfer rates which are low speed (1.5 Mbit/s) and full
speed (12 MBit/s). At this point, due to the transfer limitation, it only supports
limited types of devices like keyboards, mice, etc.

%\noindent\outline{USB2.0}\\
In 2000, the \ac{USB} 2.0~\cite{usb20} specification was released. With high speed \mbox{(480
Mbit/s)} mode introduced, printers, cameras, CD-ROM drives, and network cards are
supported in this revision. Such a high data transfer rate also give rise to the
popularity of ``flash drive'', a portable device that allows physically
transferring data around~\cite{sok}. Although various peripherals are supported
in \ac{USB} 2.0, there is no reliable way to identify the type of device. This
security flaw allows attacks like BadUSB~\cite{badusb,rubber}.

%\noindent\outline{USB3.x}\\
\ac{USB} 3.0~\cite{usb30} was introduced in 2008, with a super speed \mbox{(5 Gbit/s)} data
transfer rate. Like its predecessor, more classes of peripherals are supported
in this revision. In 2013, the \ac{USB} Type-C connector standard was introduced as a
part of \ac{USB} 3.1~\cite{usb31}, providing a unified connector type for
PowerDelivery, Thunderbolt, DisplayPort, and HDMI.  Yet no improvement of
security is introduced in 3.x revisions, meaning any device claiming
itself as a monitor can capture the video stream from the host. Exposing
such a multi-purpose connector unprotected is insecure and allows attacks similar to \tool. In 2017, \ac{USB} 3.2~\cite{usb32} was released, doubling the data
transfer rate (20 Gbit/s).

@@

Thus, to provide support towards a wider range of peripherals, a 24-pins
standard called \ac{USB} Type-C~\cite{typec} was introduced in 2013 by \ac{USB}-IF~\cite{usbif}. As it is designed to be double-sided, the number of actually usable
pins is halved. Nevertheless, this standard has largely enhanced the capability
of the \ac{USB} 3.x protocol. As presented in Figure~\ref{fig:usb_c_conn}, Type-C add
two high-speed data lanes \mbox{(TRX1 \& TRX2)} and keep the original data lane \mbox{(D+ \&
D-)}. The added lanes are used exclusively to support peripherals like
DisplayPort while the kept data lane transfers \ac{USB} packets.

### 2021-03-22 WOOT/Files/badusb.tex

%Though BadUSB devices\cite{badusb} like Rubber Ducky\cite{rubber,
%rubberducky2020} emulates as a HID device enabling various arbitrary execution
%attacks, fetching feedback from the victim is much more limited.
Based on the drawbacks introduced above, several defense mechanisms were
proposed. For example, GoodUSB~\cite{tian2015defending}
implemented an authorization procedure when a new device is plugged in and blocks other functions except authorized ones.

>>>>

%Though BadUSB devices\cite{badusb} like Rubber Ducky\cite{rubber,
%rubberducky2020} emulates as a HID device enabling various arbitrary execution
%attacks, fetching feedback from the victim is much more limited.
Based on the drawbacks introduced above, several defense mechanisms were
proposed. For example, GoodUSB~\cite{tian2015defending}
implements an authorization procedure when a new device is plugged in and blocks other functions except authorized ones.

### 2021-03-22 WOOT/Files/experiment.tex

\subsection{HID Emulator Mode}

In the experiment of HID emulator mode, we used {Lenovo Xiaoxin Pro 13
2020}, a PC with Windows 10 \mbox{(OS Build 18363.1379)} and Ubuntu 20.04.2 LTS with two \ac{USB} Type-C interfaces as the
target devices. During this experiment, \tool disguised itself as a normal keyboard and an external display as a feedback channel. When first plugged in, the victim's device raised a pop-up window asking the victim which mode should the new screen be set to. \tool immediately injected a series of mouse movements and clicks to set itself as a mirror to the primary display. Thus we had successfully completed the attack initialization. Then we tested three scripts, ranging from reverse shell backdoor to malware payload execution, all of them resulted in success.
Compared with original BadUSB attacks like Rubber Ducky, our \tool can provide attackers real-time feedback, which allows attackers to conduct more accurate attacks.
%This also warned us that we need more thorough defense against such BadUSB
%attack.

>>>>

\subsection{HID Emulation Mode}

In the experiment of HID emulation mode, we used {Lenovo Xiaoxin Pro 13
2020}, a PC with Windows 10 \mbox{(OS Build 18363.1379)} and Ubuntu 20.04.2 LTS with two \ac{USB} Type-C interfaces as the
target devices. During this experiment, \tool disguised itself as a normal keyboard and an external display as a feedback channel. When first plugged in, the victim's device raised a pop-up window asking the victim which mode should the new screen be set to. \tool immediately injected a series of mouse movements and clicks to set itself as a mirror to the primary display. Thus we had successfully completed the attack initialization. Then we tested three scripts, ranging from reverse shell backdoor to malware payload execution, all of them resulted in success.
Compared with original BadUSB attacks like Rubber Ducky, our \tool can provide attackers real-time feedback, which allows attackers to conduct more accurate attacks.
%This also warned us that we need more thorough defense against such BadUSB
%attack.

### 2021-03-22 WOOT/Files/introduction.tex

Moreover, we implemented multiple attacking
modes of \ac{USB} attacks based on our approach to verify its effectiveness,
including \ac{HID} emulator mode, video capture mode, and full control mode  To improve the
efficiency and performance of \tool, we designed a filtering algorithm to
preprocess the video data before network transmission.
We conducted
a series of experiments for each attack mode as well as for different types of
devices, including smartphones, personal computers, and tablet computers, to
validate the usability of \tool.  We also conducted a case study for attacks in
sharing power banks, one of the application scenarios of \tool.
After the validation of our attack model, we
proposed several defense methods as countermeasures, including external
hardware authorization, distrust-by-default, etc.  It is worth noting that we
designed a method, called isolated \ac{UI} rendering, to separate the user interface
into sensitive and insensitive layers.  Only the insensitive layer's content would be passed to the insecure driver and thus rendered on the external display, protecting the sensitive layer's content.
%\shuqing{Case study, evaluation, notable results.}

>>>>

Moreover, we implemented multiple attacking
modes of \ac{USB} attacks based on our approach to verify its effectiveness,
including \ac{HID} emulation mode, video capture mode, and full control mode  To improve the
efficiency and performance of \tool, we designed a filtering algorithm to
preprocess the video data before network transmission.
We conducted
a series of experiments for each attack mode as well as for different types of
devices, including smartphones, personal computers, and tablet computers, to
validate the usability of \tool.  We also conducted a case study for attacks in
sharing power banks, one of the application scenarios of \tool.
After the validation of our attack model, we
proposed several defense methods as countermeasures, including external
hardware authorization, distrust-by-default, etc.  It is worth noting that we
designed a method, called isolated \ac{UI} rendering, to separate the user interface
into sensitive and insensitive layers.  Only the insensitive layer's content would be passed to the insecure driver and thus rendered on the external display, protecting the sensitive layer's content.
%\shuqing{Case study, evaluation, notable results.}

### 2021-03-22 WOOT/Files/related_work.tex

From the hardware perspective, the BadUSB attack requires \mbox{(D+ \& D-)} pins which
are defined by the protocol to transmit data. Without these pins, data cannot 
be transferred via a \ac{USB} cable. Based
on this fact, \ac{USB} Condom~\cite{Condom} is a hardware solution to block data
channels by adding a blocker in the connector. This blocker can cut off the \mbox{(D+ \& D-)} connection while leaving the power pins intact. However, this method poses
a great challenge to the \ac{PnP} property of \ac{USB}, as once it is deployed, it
stops all \ac{USB} functions other than charging.

Under the premise of ensuring the full functionality of \ac{USB} devices, some works
improve the security while establishing connection. Windows Defender
ATP~\cite{windenfenderwhite} maintains a whitelist of \ac{USB} devices, only devices
on the whitelist are allowed to communicate with the host. This prevents all
potential attacks from untrusted devices, however, this requires users to have a
certain security awareness and technical background to maintain a valid
whitelist. For example, a naive user may add the \ac{USB} device from unknown
sources to his/her whitelist without precaution. Fortunately, some designs have been proposed to overcome
this drawback. For instance, Mohammadmoradi \emph{et al.}~\cite{mohammadmoradi2018making}\shuqing{check name} propose a strategy to generate such a
whitelist automatically. This strategy first generates a unique fingerprint for
each device based on its functionality. Then these fingerprints are used to
maintain a secure and valid whitelist of \ac{USB} devices. There is another work
mediating \ac{USB} connectivity for industrial control. TMSUI~\cite{yang2015tmsui}
relies on the rich experience of administrators to build a whitelist. However, some
modified \ac{USB} devices may hide their real functionality from the user.

To solve this flaw, GoodUSB~\cite{tian2015defending} reports the functionality claimed
by the device to the user and lets the user decide whether to authorize. When a device
is plugged in, GoodUSB blocks its functionality before enumeration until a
series of authorizations are completed. These authorizations are designed to be
performed manually, thus the malicious devices will be detected if they make a different claim 
that does not match the authorization results. 
USBeSafe\cite{usbesafe} analyses the 
feature vector constructed with the data collected from USB Request Blocks, 
detects the novel \ac{USB} packets based on the machine learning algorithm and sends 
an alert to the user before enumeration. As BadUSB attacks normally request privileged 
interfaces during enumerating, these defenses are sufficient for these attacks.
Moreover, Mueller \emph{et al.}~\cite{MuellerZN19} improve the security without changing user experience. It can detect the presence of the user, block malicious devices in the user's absence until the session is unlocked.

After the \ac{USB} enumeration and driver loading, some related works follow the \textit{defense in depth} approach to defend against BadUSB attacks.  
Neuner \emph{et al.}~\cite{neuner2018usblock}
prevent BadUSB attacks from the malicious flash drive by analyzing
the temporal characteristics of BadUSB-like attacks. This defense mechanism is
effective because the attacker cannot obtain the screen of the victim host using
BadUSB. In this case, the malicious device can only inject keystrokes in a very
short time to reduce the risk of being discovered. 
This property causes the
typing characteristics of BadUSB to be detectable. Pham \emph{et al.}~\cite{pham2010optimizing} optimize Windows security features, which can
block the execution of unsigned files and the installation of unsigned drivers
carried on portable media. Moreover, in GoodUSB, a VM is deployed in the host as
a honeypot to detect and stop malicious behavior of \ac{USB} devices.

In addition to injection attacks, data theft attack is also one of the focuses
of the academic community. As mentioned in Section~\ref{subsec:usb_attack}, there
exists an attack called \ac{JFA}~\cite{JFC} which abuses the \ac{MHL}
standard to steal the video stream from the victim. In order to mitigate this
issue, Meng et al.~\cite{meng2018252} propose a statistical model using status
like GPU/CPU usage to detect \ac{JFA} attacks.

Therefore, there exists a clear trade-off between the effectiveness and the
\ac{PnP} property. Although hardware disabling solutions like \ac{USB} Condom
achieves almost absolute security, the functionality of \ac{USB} is sacrificed.
Other solutions like whitelist are either bypassable or insufficient
under certain scenarios. Vendors may sacrifice complete security to
improve usability, which allows attackers to take advantage of.

>>>>

From the hardware perspective, the BadUSB attack requires \mbox{(D+ \& D-)} pins which
are defined by the protocol to transmit data. Without these pins, data cannot 
be transferred via a \ac{USB} cable. Based
on this fact, \ac{USB} Condom~\cite{Condom} is a hardware solution to block data
channels by adding a blocker in the connector. This blocker can cut off the \mbox{(D+ \& D-)} connection while leaving the power pins intact. However, this method poses
a great challenge to the \ac{PnP} property of \ac{USB}, as once it is deployed, it
will stop all \ac{USB} functions other than charging.

Under the premise of ensuring the full functionality of \ac{USB} devices, some works
improve the security while establishing a connection. Windows Defender
ATP~\cite{windenfenderwhite} maintains a whitelist of \ac{USB} devices, only devices
on the whitelist are allowed to communicate with the host. This prevents all
potential attacks from untrusted devices, however, this requires users to have a
certain security awareness and technical background to maintain a valid
whitelist. For example, a naive user may add the \ac{USB} device from unknown
sources to his/her whitelist without precaution. Fortunately, some designs have been proposed to overcome
this drawback. For instance, Mohammadmoradi \emph{et al.}~\cite{mohammadmoradi2018making}\shuqing{check name} propose a strategy to generate such a
whitelist automatically. This strategy first generates a unique fingerprint for
each device based on its functionality. Then these fingerprints are used to
maintain a secure and valid whitelist of \ac{USB} devices. There is another work
mediating \ac{USB} connectivity for industrial control. TMSUI~\cite{yang2015tmsui}
relies on the rich experience of administrators to build a whitelist. However, some
modified \ac{USB} devices may hide their real functionality from the user.

To solve this flaw, GoodUSB~\cite{tian2015defending} reports the functionality claimed
by the \ac{USB} device to the user and lets the user decide whether to authorize. When a device
is plugged in, GoodUSB blocks its functionality before enumeration until a
series of authorizations are completed. These authorizations are designed to be
performed manually, thus the malicious devices will be detected if they make a different claim 
that does not match the authorization results. 
USBeSafe\cite{usbesafe} analyses the 
feature vector constructed with the data collected from USB Request Blocks, 
detects the novel \ac{USB} packets based on the machine learning algorithm and sends 
an alert to the user before enumeration. As BadUSB attacks normally request privileged 
interfaces during enumerating, these defenses are sufficient for these attacks.
Moreover, Mueller \emph{et al.}~\cite{MuellerZN19} improve the security without changing the user experience. It can detect the presence of the user, block malicious devices in the user's absence until the session is unlocked.

After the \ac{USB} enumeration and driver loading, some related works follow the \textit{defense in depth} approach to defend against BadUSB attacks.  
Neuner \emph{et al.}~\cite{neuner2018usblock}
prevent BadUSB attacks from the malicious flash drive by analyzing
the temporal characteristics of BadUSB-like attacks. This defense mechanism is
effective because the attacker cannot obtain the screen of the victim's device using
BadUSB. In this case, the malicious device can only inject keystrokes in a very
short time to reduce the risk of being discovered. 
This property causes the
typing characteristics of BadUSB to be detectable. Pham \emph{et al.}~\cite{pham2010optimizing} optimize Windows security features, which can
block the execution of unsigned files and the installation of unsigned drivers
carried on portable media. Moreover, in GoodUSB, a VM is deployed in the host as
a honeypot to detect and stop malicious behavior of \ac{USB} devices.

In addition to injection attacks, data theft attack is also one of the focuses
of the academic community. As mentioned in Section~\ref{subsec:usb_attack}, there
exists an attack called \ac{JFA}~\cite{JFC} which abuses the \ac{MHL}
standard to steal the video stream from the victim's mobile phone. In order to mitigate this
issue, Meng et al.~\cite{meng2018252} propose a statistical model using status
like GPU/CPU usage to detect \ac{JFA} attacks.

Therefore, there exists a clear trade-off between the effectiveness and the
\ac{PnP} property. Although hardware disabling solutions like \ac{USB} Condom
achieve almost absolute security, the functionality of \ac{USB} is sacrificed.
Other solutions like whitelist are either bypassable or insufficient
under certain scenarios. Vendors may sacrifice complete security to
improve usability, which allows attackers to take advantage of.
